Secure Composition of Robust and Optimising Compilers

### 2023-07-09 popl/main.tex
intro feedback

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[POPL '24]{51st ACM SIGPLAN Symposium on Principles of Programming Languages}{January 17-19,
  2024}{London, UK}
%
%  Uncomment \acmBooktitle if th title of the proceedings is different
%  from ``Proceedings of ...''!
%
%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%  June 03--05, 2018, Woodstock, NY} 
\acmPrice{15.00}
\acmISBN{978-1-4503-XXXX-X/18/06}

@@

  Memory safety necessitates secrecy, since breaking it can lead to an arbitrarily large attack space, i.e., from simply reading a secret up to alteration of control--flow in order to execute arbitrary code.
  However, secrecy cannot be achieved with just memory safety alone, since, e.g., private data could be leaked by differences in execution time for varying program inputs.
  This leakage is avoidable by ensuring that programs are cryptographic constant time, i.e., different inputs do not change execution time.
  Prior work has shown that strategies to ensure cryptographic constant time can be rendered useless, since a compiler may simply optimize them away.
  This is why secure compilation is an integral ingredient to achieve security, because a secure compiler guarantees (formally) that properties that hold in the source, also hold in the target. %% mention attacker? the text describes correctness

  Unfortunately, it is an open question wether secure compilers can be engineered in a modular way.
  Ideally, compiler engineers develop two different compilation passes, one that ensures memory safety and another one that is security preserving with respect to cryptographic constant time. %, because this style of engineering is well--established for its simplifying divide--and--conquer approach as well as enhanced reusability.

  This paper answers that question positively: The composition of compilers that are secure with respect to some security--properties yields the intersection of these properties.
  This article discusses different variants of compositions of secure compilers and instantiates parts of these results in a case--study.
  To this end, it presents an optimizing, secure compilation chain that preserves both memory safety and cryptographic constant time.
%  Memory safety is achieved by composing secure compilers for both temporal and spatial memory safety, where the compiler for spatial memory safety is an instrumentation that inserts bounds--checks in front of pointer dereferences.
  %The compiler for cryptographic constant time exploits the use of a data (operand) timing independent mode and the employed optimisations of the compilation chain are constant folding and dead code elimination.
  The compositionality results enable modular proofs for the presented, secure compiler. % that is secure with respect to both memory safety and cryptographic constant time.
  Some results of this work are implemented in the Coq proof assistant.

% Secure compilers emit code that is hardened against adversarial target--level attacks, such as memory corruption or alteration of control--flow.
% It is an open question wether secure compilers compose in a well--behaved manner.
% This paper answers this question positively: The composition of compilers that are secure with respect to some security--properties yields the intersection of these properties.
% This article discusses different variants of compositions of secure compilers and instantiates parts of these results in a case--study.
% To this end, it presents an optimizing, secure compilation chain that preserves both memory safety and cryptographic constant time.
% Memory safety is achieved by composing secure compilers for both temporal and spatial memory safety, where the compiler for spatial memory safety is an instrumentation that inserts bounds--checks in front of pointer dereferences.
% The compiler for cryptographic constant time exploits the use of a data (operand) timing independent mode and the employed optimisations of the compilation chain are constant folding and dead code elimination.
% The compositionality results enable modular proofs for the whole compilation chain that is secure with respect to both memory safety and cryptographic constant time.

@@

\gls{ms} is a security property of programs.
Without memory safety, private data can easily leak or corrupt~\cite{lemay2021ccc}.
Because of this, \gls{ms} is a requirement for programs to attain an even stronger security property, namely \gls{cct}.
Programs that attain \gls{cct} are secure with respect to timing attacks, and this is the gold standard for secure applications, including cryptographic ones since timing attacks can expose secrets via a simple performance analysis~\cite{kocher1996timing}.
Timing attacks have become more and more relevant recently, especially given the era of cloud-computing~\cite{aviram2010cloudtime,kumar2019cloudsecsurvey}, where programs share the same hardware but are supposed to run in a sandboxed environment that does not leak secrets.
Unfortunately, several attacks within such a cloud environment have been demonstrated as viable strategies to extract secrets~\cite{mehmet2015getoff,flowers2022zeroday,atya2019catchme,venkatanathan2015placevul}.

Even though mitigations for such attacks exist~\cite{bond2017vale,almeida2017jasmin}, the literature shows that some compilers may optimize these mitigations away~\cite{barthe2018sec}.
This is of course not what the programmer expects and necessitates the need for a {\em secure} compiler that ensures security properties hold even when the program runs in a hostile environment, interacting with other possibly malicious programs.
Because of this hostile environment, secure compilers attain security by relying on protection mechanisms, such as ... and ...\MK{cits}.
The aim of this paper is to devise the theoretical foundation to be able to build a secure compiler that is capable of dealing with the following example.
\MPin{
	this does not really introduce the compositionality issue.
	i think the problem is the last sentence.
	try and divert the story after the 'because' line, to go into the need for composing compilers
}
\begin{example}[strncpy]\label{ex:strncpy}
Consider the program below which copies a \texttt{C}-style string \texttt{x} into \texttt{y} up to a length of \texttt{n}:

@@

For sake of simplicity, assume that the buffer \texttt{y} is at least as large as \texttt{n} as well as \texttt{x} terminated by \texttt{'\textbackslash 0'}.
Even under these assumptions, there is a subtle \gls{ms} bug in the program which can be exploited to read sensitive data:
The bounds check, i.e., \texttt{i < n}, should happen {\it prior} to the access to memory location \texttt{x[i]}.
Otherwise, if \texttt{n} is larger than the length of the array object \texttt{x} points to, the memory location past this array object is read, which is an out--of--bounds memory access.
Unfortunately, even though flipping the conditions does fix the \gls{ms} issue, it still violates \gls{cct}, so the program is susceptible to timing attacks.
This is because,\MP{this is a german comma. english does not work this way :) please erase.} the loop can terminate early: As soon as the string--\MP{only 1 dash here (also happened before but i already edited)}terminating character \texttt{'\textbackslash 0'} is encountered, the loop exits.
Thus, the program execution time is proportional to the length of the array object \texttt{x}.
The fix commonly employed in such a case is to rewrite the program to run a {\it fixed} number of iterations regardless of the length of \texttt{x} or \texttt{y}.
To this end, there must be an upper bound for the length of \texttt{x} and \texttt{y} and then the loop can just iterate until the induction variable \texttt{i} hits this upper bound.

These fixes alone are not enough to make this program robust, i.e., not violating either \gls{ms} or \gls{cct} in a malicious environment.
Consider a program context that calls \texttt{strncpy} with a very large \texttt{n}, some uninitialised memory \texttt{x}, and \texttt{y} whatever\MP{? whatever?}.
Such a function invocation leads to a series of reads from uninitialised memory, which is an immediate \gls{ms} issue.
\MPin{
	we have a very very high-level definition of MS, so the issue may not be clear
}
To mitigate these situations, it is important that \texttt{strncpy} ensures that the provided arguments satisfy the assumptions being made by the programmer, since the environment\MP{env? ctx? tetrminology} can behave in a hostile manner.
Unfortunately, these checks do not work in C\MP{why are we talking of C now?} without additional features, since there is no way to check wether\MP{please use a spell checker, this word is written 'whether'} a pointer points into a valid allocated region or not.
To enable this would require extra metadata~\cite{kwon2013lowfat,akritidis2009baggy} which must not be modifiable by a malicious context.
So, for sake of simplicity, assume the language at hand provides an operator \texttt{is\_valid(p, s, n)} that faithfully checks wether a pointer \texttt{p} points into a valid array of objects sized \texttt{s} of length \texttt{n} or not --- in constant time.
This could be an abstraction that the operating system provides.\MP{cits absolutely needed and the narrative needs to be stronger, using those cits}
With this feature at hand, it is possible to modify the original \texttt{strncpy} to be robust, given this is the compiled program without optimisations:

@@

\MPin{
	this bit should be introduced earlier
}
As stated earlier, this work wants to establish the theoretical foundation to ensure the secure compilation of programs like \texttt{strncpy} (\Cref{ex:strncpy}) to satisfy properties \gls{ms} and \gls{cct} robustly.
Secure compilation in terms of robust preservation is well-established~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}.
Unfortunately, currently, it is unclear how the theory of robust preservation supports modularity.
Missing out on modularity would mean that one has to make a secure compiler for all possible properties {\em or} for a combination of \gls{ms} and \gls{cct}.
Ideally, both for proof engineering and reusability, a secure compiler for just \gls{ms} and a secure compiler for just \gls{cct} can compose in some way to get a secure compiler for the combination of the two.
This also reflects real--world compilation chains, where several compilation passes perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source--code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfills a property of interest.

% keyword: pass
Thus, this paper introduces a framework for reasoning about the composition of secure and optimising compiler passes and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
To this end, this paper first discusses how to compose security properties, such as temporal and spatial memory safety into general memory safety, and cryptographic constant time.
Then, this paper defines several secure compiler passes, where each is either preserving a different security property (e.g., temporal or spatial memory safety) or performing a security-preserving optimisation, (e.g., applying constant-folding or dead-code elimination).
% keyword- end to end, composition
Finally, this paper shows how to compose these secure compiler passes into a multi-pass compilation chain which is proven to provide end-to-end preservation of general memory safety and cryptographic constant time.
Crucially, this end-to-end security preservation is obtained by having each compilation pass (i) preserve a sub-part of the overall security property individually and then (ii) compose each preserved security property as dictated by the framework of this paper.
The results showcase how the framework allows the kind of formal security reasoning that compiler writers already want (and do), obtaining precise, compositional security reasoning while providing minimal (and modular) proof effort.
\MPin{
	last line: connect with the many citations of the previous paragraph
}

In summary, this paper makes the following contributions:
\begin{itemize}
  \item %
        %\Cref{sec:compprop} presents a formal framework to reason about compositions of security properties and demonstrate its usage with concrete examples:
        First, this paper studies the formalisation and the composition of security properties (\Cref{sec:compprop}) by focussing on those properties that are of interest for real-world compiler writers (as identified by the plethora of work enforcing such properties individually~\cite{akritidis2009baggy,nagarakatte2009soft,nagarakatte2010cets,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact}).
        % 
        Starting from ways to formalise and enforce those properties individually, this paper shows how to compose their formalisation. % and their composition. %% MK: compose their composition??
        % 
        This paper then showcases the benefits of this composition by composing \gls{tms} and \gls{sms} into \gls{ms}, and ultimately adding \gls{cct} to the composition, yielding the the conjunction of \gls{ms} and \gls{cct}.
        % 
        The resulting security property is the golden standard of safety properties that secure cryptographic code should have~\cite{lemay2021ccc}\MK{more cits?}.
%        \MPin{
%        	conclude that did this prop in a single pass, we show how to do it compositionally
%        }

  \item %
        %\Cref{sec:compcomp} studies different forms of secure compiler compositions.
        This paper takes the secure compilation framework of~\citep{abate2019jour} and extends it to reason about the security of all different known forms of compiler composition (\Cref{sec:compcomp}).
        %Concretely, it investigates upper and lower compositions, as well as the sequential composition.
        For this, this paper studies the sequential composition of compilers, as typically done when engineering compilation chains, as well as an upper and a lower composition that discuss multi--in- and multi--output compilers.\MP{upper and lower?}
        % 
        This paper proves that starting from two compilers that preserve two properties, their composition preserves the intersection of those properties.
        For example, a compiler that is secure with respect to \gls{ms} can be composed with a compiler that is secure with respect to \gls{cct}, resulting in a compiler that is secure with respect to the composition of \gls{ms} and \gls{cct}.
%       \MPin{
%      		find way to connect to previous contrib
%       }
        % 
        Finally, this paper provides an important corollary that demonstrates that the order of composition of sequential compiler passes is irrelevant for the resulting security.
        % 
        This corollary is crucial for reordering compiler optimisation passes as well as secure compilation passes and thus generating secure and efficient code.

  \item %
        This paper presents a case-study showcasing the conjunction of the previous contributions (\Cref{sec:casestud:defs,sec:casestud:rtp}).
        To this end, it presents a compilation chain consisting of several passes.
        The chain ultimately preserves a combination of \gls{ms} and \gls{cct} by means of composing the individual, secure passes concerning \gls{tms}, \gls{sms}, and a stricter variant of \gls{cct}, respectively.
        Furthermore, the chain includes two passes for \gls{ms} that are optimising: One performs a simple \gls{dce} and the other \gls{cf}.
        The formalisation of this case study showcases the power of the presented framework.
        Namely, the divide--and--conquer approach is a viable strategy even for engineering secure compilers.
%        This demonstrates that the usage of the presented framework is both simpler and more expressive compared to one monolithic security proof of the whole compilation chain.

@@

This paper employs the security model where programs are written in a language whose semantics emits events $\event$.
Events include security--relevant actions (e.g., reading from and writing to memory) and the unobservable event $\emptyevent$.
As programs execute, their emitted events are concatenated in traces $\trace$, which serve as the description of the behaviour of a program.%
\footnote{
Throughout the paper, we indicate sequences with an overbar (i.e., $\trace$), we denote empty sequences with $\hole{\cdot}$ and concatenate sequences $\trace[_{1}],\trace[_{2}]$ as $\trace[_{1}]\cdot\trace[_{2}]$.
Prepending elements to sequences uses the same notation: $\event\cdot\trace$.
}

Properties $\pi$ are sets of traces of admissible program behaviours, ascribing what said property considers valid~\cite{clarkson2008hyper}.
The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety nor liveness~\cite{clarkson2008hyper}.
A class is simply a set of properties and for the class of safety properties, it is decidable wether a trace satisfies a safety property with just a finite trace prefix.
As an example, consider a trace describing an interaction with a memory where the deallocation of an address $\loc$ precedes a read at that address in memory: $\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$
This program behaviour is insecure with respect to a canonical notion of (temporal) memory safety dictating no use-after-frees of pointers~\cite{nagarakatte2010cets}, because it reads from a memory location that was freed already.
The previous finite trace prefix is enough to decide that the trace does not satisfy temporal memory safety and there is no way to append events to this prefix which would result in the trace being admissible.
In the following, the notation of an execution of a whole program $\wholeprogvar$ that terminates in state $\runtimetermvar$ according to the language semantics and produces trace $\trace$ is $\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$.
With this, the definition of property satisfaction is as follows:
\bul{whole programs $\wholeprogvar$ satisfy a property $\pi$} iff \iul{$\wholeprogvar$ yields a trace $\trace$} such that \oul{$\trace$ satisifies $\pi$} (\Cref{def:propsat}).

@@

\begin{example}[Double Free in Bluetooth Subsystem]
  Consider \texttt{CVE-2021-3564}~\cite{doublefree-bluetooth}, one of many submissions for a double--free vulnerability.
  The bug arises due to a race condition where the context--level function \texttt{hci\_cmd\_work} was not expected to behave maliciously, since it resides in the same source--code repository where the vulnerability occurs.
  Nevertheless, the component--level code of \texttt{hci\_dev\_do\_open} is linked with \texttt{hci\_cmd\_work} and does not atomically check wether the pointer has been freed already:
  Therefore, \texttt{hci\_dev\_do\_open} does not satisfy the no--double--frees property robustly, since there is an implementation for \texttt{hci\_cmd\_work} that results the whole program to violate that property.
\end{example}

@@

A {\em compiler} ($\cc{\src{L}}{\trg{L}}$) translates syntactic descriptions of programs from a {\em source} ($\src{L}$) into a {\em target} ($\trg{L}$) programming language.
%For the syntactic description of programs written in $\src{L}$, this paper writes $\src{\progvar}$, while for ones written in $\trg{L}$, this paper writes $\trg{\progvar}$.%\MP{not a fan of this sentence}
%This paper writes $\src{\progvar}$ and $\trg{\progvar}$ for programs written in $\src{L}$ and $\trg{L}$, respectively.
%      MK: I don't think we need to mention this at all
This translation is considered {\em correct} if it is semantics--preserving.
That is, for a whole program $\src{\wholeprogvar}$, the compiler should relate the $\src{L}$ semantics of $\src{\wholeprogvar}$ with the semantics of $\trg{T}$ in such a way that they are ,,compatible''.
Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey}, and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
To define when a compiler is secure, this paper uses the robust compilation framework~\cite{abate2019jour}, which we summarize in the following definition.

@@

\begin{gather*}
  \begin{aligned}
  \mi{(Security\ Tag)}~\sigma\bnfdef&\ \lock \mid \unlock\hspace{0.5cm}
  \mi{(Control\ Tag)}~\sandboxtag\bnfdef \ctx\mid\comp\hspace{0.5cm}
  \mi{(Event)}~\event\bnfdef\ \emptyevent \mid \lightning \mid \event[_{b}];\sandboxtag;\sigma \\
  \mi{(Pre\text{--}event)}&~\event[_{b}]\bnfdef\ \ev{Alloc\ \loc\ n} \mid \ev{Dealloc\ \loc} \mid \ev{Use\ \loc\ n} \mid \ev{Branch\ n} \mid \ev{Binop\ n} \\
  \end{aligned}
\end{gather*}

>>>>

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[POPL '24]{51st ACM SIGPLAN Symposium on Principles of Programming Languages}{January 17-19,
  2024}{London, UK}
%
%  Uncomment \acmBooktitle if th title of the proceedings is different
%  from ``Proceedings of ...''!
%
%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%  June 03-05, 2018, Woodstock, NY}
\acmPrice{15.00}
\acmISBN{978-1-4503-XXXX-X/18/06}

@@

  Memory safety necessitates secrecy, since breaking it can lead to an arbitrarily large attack space, i.e., from simply reading a secret up to alteration of control-flow in order to execute arbitrary code.
  However, secrecy cannot be achieved with just memory safety alone, since, e.g., private data could be leaked by differences in execution time for varying program inputs.
  This leakage is avoidable by ensuring that programs are cryptographic constant time, i.e., different inputs do not change execution time.
  Prior work has shown that strategies to ensure cryptographic constant time can be rendered useless, since a compiler may simply optimize them away.
  This is why secure compilation is an integral ingredient to achieve security, because a secure compiler guarantees (formally) that properties that hold in the source, also hold in the target. %% mention attacker? the text describes correctness

  Unfortunately, it is an open question whether secure compilers can be engineered in a modular way.
  Ideally, compiler engineers develop two different compilation passes, one that ensures memory safety and another one that is security preserving with respect to cryptographic constant time. %, because this style of engineering is well-established for its simplifying divide-and-conquer approach as well as enhanced reusability.

  This paper answers that question positively: The composition of compilers that are secure with respect to some security-properties yields the intersection of these properties.
  This article discusses different variants of compositions of secure compilers and instantiates parts of these results in a case-study.
  To this end, it presents an optimizing, secure compilation chain that preserves both memory safety and cryptographic constant time.
%  Memory safety is achieved by composing secure compilers for both temporal and spatial memory safety, where the compiler for spatial memory safety is an instrumentation that inserts bounds-checks in front of pointer dereferences.
  %The compiler for cryptographic constant time exploits the use of a data (operand) timing independent mode and the employed optimisations of the compilation chain are constant folding and dead code elimination.
  The compositionality results enable modular proofs for the presented, secure compiler. % that is secure with respect to both memory safety and cryptographic constant time.
  Some results of this work are implemented in the Coq proof assistant.

% Secure compilers emit code that is hardened against adversarial target-level attacks, such as memory corruption or alteration of control-flow.
% It is an open question whether secure compilers compose in a well-behaved manner.
% This paper answers this question positively: The composition of compilers that are secure with respect to some security-properties yields the intersection of these properties.
% This article discusses different variants of compositions of secure compilers and instantiates parts of these results in a case-study.
% To this end, it presents an optimizing, secure compilation chain that preserves both memory safety and cryptographic constant time.
% Memory safety is achieved by composing secure compilers for both temporal and spatial memory safety, where the compiler for spatial memory safety is an instrumentation that inserts bounds-checks in front of pointer dereferences.
% The compiler for cryptographic constant time exploits the use of a data (operand) timing independent mode and the employed optimisations of the compilation chain are constant folding and dead code elimination.
% The compositionality results enable modular proofs for the whole compilation chain that is secure with respect to both memory safety and cryptographic constant time.

@@

\gls{ms} is a security property of programs that, e.g., ensures that there are no out-of-bounds accesses as well as no reads or writes to uninitialized memory.
Without memory safety, private data can easily leak or corrupt~\cite{lemay2021ccc}.
Because of this, \gls{ms} is a requirement for programs to attain an even stronger security property, namely \gls{cct}.
Programs that attain \gls{cct} are secure with respect to timing attacks, and this is the gold standard for secure applications, including cryptographic ones since timing attacks can expose secrets via a simple performance analysis~\cite{kocher1996timing}.
Timing attacks have become more and more relevant recently, especially given the era of cloud-computing~\cite{aviram2010cloudtime,kumar2019cloudsecsurvey}, where programs share the same hardware but are supposed to run in a sandboxed environment that does not leak secrets.
Unfortunately, several attacks within such a cloud environment have been demonstrated as viable strategies to extract secrets~\cite{mehmet2015getoff,flowers2022zeroday,atya2019catchme,venkatanathan2015placevul}.

Even though mitigations for such attacks exist~\cite{bond2017vale,almeida2017jasmin}, the literature shows that some compilers may optimize these mitigations away~\cite{barthe2018sec}.
This is of course not what the programmer expects and necessitates the need for a {\em secure} compiler that ensures security properties hold even when the program runs in a hostile context, i.e., where it interacts with other possibly malicious programs.
Secure compilers can be described formally with the theory on {\em robust preservation}~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}, which takes this interaction between components and potentially malicious contexts into account.
Unfortunately, it is unclear if and how robust preservation supports modularity.
Modularity is an inherent part of software engineering and real-world compilers are no exception to that~\cite{lattner2004llvm,googlev8}.
Compilation passes are typically developed as a standalone module that can plugged into a compilation pipeline.
Different kinds of such passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfills a property of interest.
Without a modular framework, there is no hope to enable this separation of concerns, where, e.g., a secure pass that ensures \gls{ms} can be developed independently of a secure pass for \gls{cct}.
%% MK: no idea how/where to fit this in or if this is even necessary
%Because of the potentially hostile context, secure compilers attain security by relying on protection mechanisms, such as ... and ...\MK{cits}.
The aim of this paper is to devise the theoretical foundation to be able to modularily build a secure compiler that is capable of dealing with the following example.
%\MPin{ MK: hope it's solved now
%	this does not really introduce the compositionality issue.
%	i think the problem is the last sentence.
%	try and divert the story after the 'because' line, to go into the need for composing compilers
%}
\begin{example}[strncpy]\label{ex:strncpy}
Consider the program below which copies a \texttt{C}-style string \texttt{x} into \texttt{y} up to a length of \texttt{n}:

@@

For sake of simplicity, assume that the buffer \texttt{y} is at least as large as \texttt{n} as well as \texttt{x} terminated by \texttt{'\textbackslash 0'}.
Even under these assumptions, there is a subtle \gls{ms} bug in the program which can be exploited to read sensitive data:
The bounds check, i.e., \texttt{i < n}, should happen {\it prior} to the access to memory location \texttt{x[i]}.
Otherwise, if \texttt{n} is larger than the length of the array object \texttt{x} points to, the memory location past this array object is read, which is an out-of-bounds memory access.
Unfortunately, even though flipping the conditions does fix the \gls{ms} issue, it still violates \gls{cct}, so the program is susceptible to timing attacks.
This is because the loop can terminate early: As soon as the string-terminating character \texttt{'\textbackslash 0'} is encountered, the loop exits.
Thus, the program execution time is proportional to the length of the array object \texttt{x}.
The fix commonly employed in such a case is to rewrite the program to run a {\it fixed} number of iterations regardless of the length of \texttt{x} or \texttt{y}.
To this end, there must be an upper bound for the length of \texttt{x} and \texttt{y} and then the loop can just iterate until the induction variable \texttt{i} hits this upper bound.

These fixes alone are not enough to make this program robust, i.e., not violating either \gls{ms} or \gls{cct} in a potentially malicious context.
Consider a program context that calls \texttt{strncpy} with a very large \texttt{n}, some uninitialised memory \texttt{x}, and \texttt{y} does not matter for the sake of this example.
While the context seems conceptually simple, uninitialized memory can appear in programs very subtly and invoking \texttt{strncpy} now leads to a series of reads from uninitialised memory, which is an immediate \gls{ms} issue that can have devastating real-world implications~\cite{uninit-0,uninit-1,uninit-2,uninit-3,uninit-4}.
To mitigate these situations, it is important that \texttt{strncpy} ensures that the provided arguments satisfy the assumptions being made by the maintainer of \texttt{strncpy}, since the calling context can behave in a hostile manner.
Unfortunately, these checks do not work in C\MP{why are we talking of C now?} without additional features, since there is no way to check whether a pointer points into a valid allocated region or not.
To enable this would require extra metadata~\cite{kwon2013lowfat,akritidis2009baggy} which must not be modifiable by a malicious context.
So, for sake of simplicity, assume the language at hand provides an operator \texttt{is\_valid(p, s, n)} that faithfully checks whether a pointer \texttt{p} points into a valid array of objects sized \texttt{s} of length \texttt{n} or not -- in constant time.
This could be an abstraction that the operating system provides.\MP{cits absolutely needed and the narrative needs to be stronger, using those cits}
With this feature at hand, it is possible to modify the original \texttt{strncpy} to be robust, given this is the compiled program without optimisations:

@@

This paper introduces a framework for reasoning about the composition of secure and optimising compiler passes and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
To this end, this paper first discusses how to compose security properties, such as temporal and spatial memory safety into general memory safety, and cryptographic constant time.
Then, this paper defines several secure compiler passes, where each is either preserving a different security property (e.g., temporal or spatial memory safety) or performing a security-preserving optimisation, (e.g., applying constant-folding or dead-code elimination).
Finally, this paper shows how to compose these secure compiler passes into a multi-pass compilation chain which is proven to provide end-to-end preservation of general memory safety and cryptographic constant time.
Crucially, this end-to-end security preservation is obtained by having each compilation pass (i) preserve a sub-part of the overall security property individually and then (ii) compose each preserved security property as dictated by the framework of this paper.
The results showcase how the framework allows the kind of formal security reasoning that compiler writers already want (and do), obtaining precise, compositional security reasoning while providing minimal (and modular) proof effort.
%\MPin{
%	last line: connect with the many citations of the previous paragraph
%}

In summary, this paper makes the following contributions:
\begin{itemize}
  \item %
        %\Cref{sec:compprop} presents a formal framework to reason about compositions of security properties and demonstrate its usage with concrete examples:
        First, this paper studies the formalisation and the composition of security properties (\Cref{sec:compprop}) by focussing on those properties that are of interest for real-world compiler writers (as identified by the plethora of work enforcing such properties individually~\cite{akritidis2009baggy,nagarakatte2009soft,nagarakatte2010cets,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact}).
        % 
        Starting from ways to formalise and enforce those properties individually, this paper shows how to compose their formalisation. % and their composition. %% MK: compose their composition??
        % 
        This paper then showcases the benefits of this composition by composing \gls{tms} and \gls{sms} into \gls{ms}, and ultimately adding \gls{cct} to the composition, yielding the the conjunction of \gls{ms} and \gls{cct}.
        % 
        The resulting security property is the golden standard of safety properties that secure cryptographic code should have~\cite{lemay2021ccc}.
%        \MPin{
%        	conclude that did this prop in a single pass, we show how to do it compositionally
%        }

  \item %
        %\Cref{sec:compcomp} studies different forms of secure compiler compositions.
        This paper takes the secure compilation framework of~\citep{abate2019jour} and extends it to reason about the security of all different known forms of compiler composition (\Cref{sec:compcomp}).
        %Concretely, it investigates upper and lower compositions, as well as the sequential composition.
        For this, this paper studies compilers with multiple inputs or multiple outputs as well as the sequential composition of compilers, as typically done when engineering compilation chains.
        %
        This paper proves that starting from two compilers that preserve two properties, their composition preserves the intersection of those properties.
        For example, a compiler that is secure with respect to \gls{ms} can be composed with a compiler that is secure with respect to \gls{cct}, resulting in a compiler that is secure with respect to the composition of \gls{ms} and \gls{cct}.
%       \MPin{
%      		find way to connect to previous contrib
%       }
        % 
        Finally, this paper provides an important corollary that demonstrates that the order of composition of sequential compiler passes is irrelevant for the resulting security.
        % 
        This corollary is crucial for reordering compiler optimisation passes as well as secure compilation passes and thus generating secure and efficient code.

  \item %
        This paper presents a case-study showcasing the conjunction of the previous contributions (\Cref{sec:casestud:defs,sec:casestud:rtp}).
        To this end, it presents a compilation chain consisting of several passes.
        The chain ultimately preserves a combination of \gls{ms} and \gls{cct} by means of composing the individual, secure passes concerning \gls{tms}, \gls{sms}, and a stricter variant of \gls{cct}, respectively.
        Furthermore, the chain includes two passes for \gls{ms} that are optimising: One performs a simple \gls{dce} and the other \gls{cf}.
        The formalisation of this case study showcases the power of the presented framework.
        Namely, the divide-and-conquer approach is a viable strategy even for engineering secure compilers.
%        This demonstrates that the usage of the presented framework is both simpler and more expressive compared to one monolithic security proof of the whole compilation chain.

@@

This paper employs the security model where programs are written in a language whose semantics emits events $\event$.
Events include security-relevant actions (e.g., reading from and writing to memory) and the unobservable event $\emptyevent$.
As programs execute, their emitted events are concatenated in traces $\trace$, which serve as the description of the behaviour of a program.%
\footnote{
Throughout the paper, we indicate sequences with an overbar (i.e., $\trace$), we denote empty sequences with $\hole{\cdot}$ and concatenate sequences $\trace[_{1}],\trace[_{2}]$ as $\trace[_{1}]\cdot\trace[_{2}]$.
Prepending elements to sequences uses the same notation: $\event\cdot\trace$.
}

Properties $\pi$ are sets of traces of admissible program behaviours, ascribing what said property considers valid~\cite{clarkson2008hyper}.
The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety nor liveness~\cite{clarkson2008hyper}.
A class is simply a set of properties and for the class of safety properties, it is decidable whether a trace satisfies a safety property with just a finite trace prefix.
As an example, consider a trace describing an interaction with a memory where the deallocation of an address $\loc$ precedes a read at that address in memory: $\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$
This program behaviour is insecure with respect to a canonical notion of (temporal) memory safety dictating no use-after-frees of pointers~\cite{nagarakatte2010cets}, because it reads from a memory location that was freed already.
The previous finite trace prefix is enough to decide that the trace does not satisfy temporal memory safety and there is no way to append events to this prefix which would result in the trace being admissible.
In the following, the notation of an execution of a whole program $\wholeprogvar$ that terminates in state $\runtimetermvar$ according to the language semantics and produces trace $\trace$ is $\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$.
With this, the definition of property satisfaction is as follows:
\bul{whole programs $\wholeprogvar$ satisfy a property $\pi$} iff \iul{$\wholeprogvar$ yields a trace $\trace$} such that \oul{$\trace$ satisifies $\pi$} (\Cref{def:propsat}).

@@

\begin{example}[Double Free in Bluetooth Subsystem]
  Consider \texttt{CVE-2021-3564}~\cite{doublefree-bluetooth}, one of many submissions for a double-free vulnerability.
  The bug arises due to a race condition where the context-level function \texttt{hci\_cmd\_work} was not expected to behave maliciously, since it resides in the same source-code repository where the vulnerability occurs.
  Nevertheless, the component-level code of \texttt{hci\_dev\_do\_open} is linked with \texttt{hci\_cmd\_work} and does not atomically check whether the pointer has been freed already:
  Therefore, \texttt{hci\_dev\_do\_open} does not satisfy the no-double-frees property robustly, since there is an implementation for \texttt{hci\_cmd\_work} that results the whole program to violate that property.
\end{example}

@@

A {\em compiler} ($\cc{\src{L}}{\trg{L}}$) translates syntactic descriptions of programs from a {\em source} ($\src{L}$) into a {\em target} ($\trg{L}$) programming language.
%For the syntactic description of programs written in $\src{L}$, this paper writes $\src{\progvar}$, while for ones written in $\trg{L}$, this paper writes $\trg{\progvar}$.%\MP{not a fan of this sentence}
%This paper writes $\src{\progvar}$ and $\trg{\progvar}$ for programs written in $\src{L}$ and $\trg{L}$, respectively.
%      MK: I don't think we need to mention this at all
This translation is considered {\em correct} if it is semantics-preserving.
That is, for a whole program $\src{\wholeprogvar}$, the compiler should relate the $\src{L}$ semantics of $\src{\wholeprogvar}$ with the semantics of $\trg{T}$ in such a way that they are ,,compatible''.
Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey}, and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
To define when a compiler is secure, this paper uses the robust compilation framework~\cite{abate2019jour}, which we summarize in the following definition.

@@

\begin{gather*}
  \begin{aligned}
  \mi{(Security\ Tag)}~\sigma\bnfdef&\ \lock \mid \unlock\hspace{0.5cm}
  \mi{(Control\ Tag)}~\sandboxtag\bnfdef \ctx\mid\comp\hspace{0.5cm}
  \mi{(Event)}~\event\bnfdef\ \emptyevent \mid \lightning \mid \event[_{b}];\sandboxtag;\sigma \\
  \mi{(Pre\text{-}event)}&~\event[_{b}]\bnfdef\ \ev{Alloc\ \loc\ n} \mid \ev{Dealloc\ \loc} \mid \ev{Use\ \loc\ n} \mid \ev{Branch\ n} \mid \ev{Binop\ n} \\
  \end{aligned}
\end{gather*}
